{RandomRationalPoints} package for \emph{Macaulay2}

### 2021-01-20 FastLinAlg/Paper/RandomRationalPoints.tex
Updated the paper to make it shorter.

\documentclass[11pt]{amsart}
\usepackage{calc,amssymb,amsthm,amsmath,fullpage}
% \usepackage{mathtools}
\RequirePackage[dvipsnames,usenames, table]{xcolor}
\usepackage{hyperref}
\usepackage[capitalize]{cleveref}
\hypersetup{
  bookmarks,
  bookmarksdepth=3,
  bookmarksopen,
  bookmarksnumbered,
  pdfstartview=FitH,
  colorlinks,backref,hyperindex,
  linkcolor=Sepia,
  anchorcolor=BurntOrange,
  citecolor=MidnightBlue,
  citecolor=OliveGreen,
  filecolor=BlueViolet,
  menucolor=Yellow,
  urlcolor=OliveGreen
}
\usepackage{xypic}
\usepackage{alltt}
\usepackage{multicol}  
% \usepackage{etex}
\usepackage{xspace}
\usepackage{rotating}
\interfootnotelinepenalty=100000
\theoremstyle{definition}
\newtheorem{example}{Example}[section]

% \usepackage{mabliautoref}
% \usepackage{colonequals}
% \frenchspacing
% \input{kmacros3.sty}
% \usepackage{stmaryrd}
% 
\usepackage{verbatim}
\usepackage{enumerate}
\begin{document}
\title{{RandomRationalPoints} package for \emph{Macaulay2}}
\author{Sankhaneel Bisui}
\email{sbisui@tulane.edu}
\address{Tulane University}
\author{Zhan Jiang}
\email{zoeng@umich.edu}
\address{University of Michigan}
\author{Sarasij Maitra}
\email{sm3vg@virginia.edu}
\address{University of Virginia}
\author{Thai Nguyen}
\email{tnguyen11@tulane.edu}
\address{Tulane University}
\author{Karl Schwede}
\email{schwede@math.utah.edu}
\address{Department of Mathematics, University of Utah, 155 S 1400 E Room 233, Salt Lake City, UT, 84112}
\date{\today}

\begin{abstract}
  In this article, we present {\tt RandomRationalPoints}, a package in \emph{Macaulay2} designed mainly to identify rational and geometric points in a variety over a finite field. We provide different strategies like linear intersection and projection to obtain such points. We also present methods to obtain non-vanishing minors of some given size in a given matrix, by evaluating the matrix at a point.  This, in conjunction with the {\tt FastLinAlg} package, should aid in determining properties of the singular locus.
\end{abstract}

@@

There is an existing package called {\tt RandomPoints}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one point quickly, even if it is not rational.  
We provide some functions that apply different strategies to generate random rational points for the given variety. We also provide functions that will expedite the process of determining properties of the singular locus of $X$ when combined with the functions provided in the package {\tt FastLinAlg}.

We provide the following core functions:
\begin{itemize}
\item {\tt projectionToHypersurface} and {\tt genericProjection}: These functions provide customizable projection. (\Cref{projectionfunctions}) 	
\item {\tt randomPoints}:  This tries to find a point in the vanishing set of an ideal. (\Cref{randomPoints})
\item {\tt findANonZeroMinor}:  This finds a submatrix of a given matrix that is nonsingular at a point of a given ideal. (\Cref{findANonZeroMinor})          
\item {\tt extendIdealByNonZeroMinor}:  This extends an ideal by a minor produced by \\{\tt findANonZeroMinor}. (\Cref{extendIdealByNonZeroMinor})	
\end{itemize}

All polynomial rings considered here will be over finite fields. We first briefly describe some helper functions that will be used extensively in the core functions above. In the subsequent sections, we explain briefly the core functions. We shall also briefly mention the strategies that we have implemented in the execution of both helper and core functions.

@@

\section{A helper function: randomCoordinateChange}{\label{helper}}

% \item {\tt Codimension}: It takes input an integer. Various functions need to know the codimension/height of the scheme/ideal it is working with.  Setting this to be an integer will tell the function not to compute the codimension and to use this value instead.  The default value is {\tt null}, in which case the function will compute the codimension.  
%%%   
%%   Codimension is not a function, it is an option.
%%%   
\subsection*{\tt randomCoordinateChange:} This function takes a polynomial ring as an input and outputs the coordinate change map.  
i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.

In some applications, a full change of coordinates is not desired, as it might cause code to run very slowly.  A binomial change of coordinates might be preferred, or we might only replace some monomials by other monomials.  
Refinements of {\tt randomCoordinateChange} are controlled with the following options.

\begin{itemize}
\item {\tt Replacement}: Setting {\tt Replacement => Full} will mean that coordinates are replaced by a general degree $1$ form.  If {\tt Replacement => Binomial}, the coordinates are only changed to binomials, which can be much faster for certain applications. If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
\item {\tt MaxCoordinatesToReplace}: The user can specify that only a specified number of coordinates should be non-monomial (assuming {\tt Homogeneous => true}).  
  This option is passed to {\tt randomCoordinateChange} by other functions that call it. 		
\item {\tt Homogeneous}:  Setting {\tt Homogeneous => false} will cause degree zero terms to be added to modified coordinates (including monomial coordinates).		
\end{itemize}

\vspace{1em}

\begin{example}~~

  {{\small\color{blue}
\begin{verbatim}
	i3 : R = ZZ/11[x, y, z];

	o7 = | y x+4z z |
\end{verbatim}}
    }}

\end{example}
\section{genericProjection, projectionToHypersurface}\label{projectionfunctions}

% \begin{enumerate}
\subsection{genericProjection} 
This function finds a random (somewhat, depending on options) generic projection of the ring or ideal.
The typical usages are as follows: 
\vspace{1em}

\vspace{1em}
\noindent where 
$I$ is an ideal 
in a polynomial ring, 
$R$ can denote a quotient of a polynomial ring and 
$n\in \mathbb{Z}$ is
an integer specifying how many dimensions to drop.  Note that this function makes no attempt to verify that the projection is actually generic with respect to the ideal.

\begin{example}	
  ~~

o3 : Sequence
\end{verbatim}}
    }}
\end{example}

\begin{example}~~

o3 : Sequence
\end{verbatim}}
    }}
\end{example}

This method works by calling {\tt randomCoordinateChange} (see \Cref{helper}) before dropping some variables.  It passes the options {\tt Replacement}, {\tt MaxCoordinatesToReplace}, {\tt Homogeneous} to that function.

The user can also specify the same options as before to be passed on to the {\tt randomCoordinateChange} function.

\newpage

\begin{example}~~

o4 : Sequence
\end{verbatim}}
    }}
\end{example}

\vspace{1em}
-- {\tt projectionToHypersurface I},

-- {\tt projectionToHypersurface R}

\vspace{1em}
\noindent where $I$ is an ideal in a polynomial ring, $R$ is a quotient of a polynomial ring. The output is
a list with two entries, the generic projection map, and the ideal if {\tt I} was provided, or the ring if {\tt R} was provided.

\vspace{1em}
\begin{example}~~

	               3       2     3            2        2       2       2  
	o3 = ideal (- x  - 2x*y  - 2y  - x*y*z + y z - 2x*z  + 2y*z  - 2, x  -
	-------------------------------------------------------------------
	        2
	2x*y - y  + x*z - y*z)

	            ZZ                             4    3      2 2     4
	o4 = (map(R,--[y, z],{x + 2y, 2y}), ideal(y  + y z - 2y z  - 2z  - y -
	            5
	-------------------------------------------------------------------
	z))

	o5 : Sequence

\end{verbatim}}
    }}
\end{example}

% \end{enumerate}
\section{randomPoints}\label{randomPoints}
{\tt randomPoints} is a function to find rational or geometric points in a variety. The typical usages are as follows:

\vspace{1em}
-- {\tt randomPoints(I)},

@@

\vspace{1em}
\noindent where $n$ is a positive integer denoting the number of desired points, and 
$I$ is an ideal inside a polynomial ring.  If {$ n$} is omitted, then it is assumed to be 1.

@@

\begin{description}		
\item[\tt Strategy => $\bullet$]   Here the $\bullet$ can be {\tt Default, BruteForce, LinearIntersection, \\ GenericProjection} or {\tt HybridProjectionIntersection}.

  \begin{itemize}
  \item {\tt Default} performs a sequence of the different strategies below, aimed at finding a point quickly.    
  \item {\tt BruteForce} simply tries random points and sees if they are on the variety.

@@

  \begin{example}\label{BruteForce}
    ~~

			i2 : R = ZZ/101[x, y, z];

@@

			o6 : List

\end{verbatim}}
      }}
  \end{example}

\item[\tt ProjectionAttempts => ZZ]

@@

  \vspace{1em}
\item[\tt MaxCoordinatesToReplace => ZZ] (see \Cref{helper})
  \vspace{1em}
\item[\tt Codimension => ZZ] (see \Cref{helper})

  \vspace{1em}
\item[\tt ExtendField => Boolean]

@@

  \vspace{1em}
\item[\tt IntersectionAttempts => ZZ]

@@

  \vspace{1em}
\item[\tt PointCheckAttempts => ZZ]

  When calling {\tt randomPoints}, and functions that call it, with a {\tt BruteForce} strategy or {\tt GenericProjection} strategy, this denotes the number of trials for brute force point checking.

  \begin{example}

    We re-compute \cref{BruteForce} this time specifying more attempts.

    {{\small\color{blue}
\begin{verbatim}
	i1: R = ZZ/101[x, y, z];

	i2 : J = ideal "x3 + y2 + 1, z3 - x2 - y2 + 2";

	o2 : Ideal of R

	i3 : randomPoints(J, Strategy => BruteForce, PointCheckAttempts => 20000)
		-- used 0.84679 seconds

	o3 = {{-39, -43, 28}}

	o3 : List
\end{verbatim}}
      }}
  \end{example}

  \vspace{1em}
\item[\tt MaxCoordinatesToTrivialize]

@@

  \vspace{1em}
\item[\tt NumThreadsToUse => ZZ]

  When calling {\tt randomPoints}, and functions that call it, with a {\tt BruteForce} strategy, this denotes the number of threads to use in brute force point checking.
\end{description}
% Outputs
% :List
% a list of points in the variety with possible repetitions.
% Description
% Text  
% Gives at most $n$ many point in a variety $V(I)$. 
% Example
% R = ZZ/5[t_1..t_3];
% I = ideal(t_1,t_2+t_3);
% randomPoints(3, I)
% randomPoints(4, I, Strategy => Default)
% randomPoints(4, I, Strategy => GenericProjection)
% randomPoints(4, I, Strategy => LinearIntersection)

@@

When working over very small fields, frequently {\tt BruteForce} is most efficient.  This is not surprising as there may not be many points to check, and a surprisingly large percentage of points can be rational in hypersurfaces and complete intersections.  However, if the field size is larger, {\tt BruteForce} will perform poorly.  Other strategies work differently on different examples, and even the same strategy can sometimes work very quickly even if it typically works very slowly.

~

\begin{example}
  ~~

			i2 : R = ZZ/7[x_1..x_10];

@@

			i6 : S = ZZ/211[x_1..x_10];

@@

			o11 = {{31, -15, 53, 41, 87, 9, -21, -43, -12, 85}}

			o11 : List

			i12 : time randomPoints(J, Strategy => LinearIntersection, ExtendField => true)
     		-- used 0.0612306 seconds

			               5      4      3      2                                      5  
			o12 = {{54, 14a  - 48a  - 56a  + 56a  - 13a + 48, -101, 56, -42, -27, - 26a  -
				-------------------------------------------------------------------------
				   4      3      2
				43a  - 70a  - 96a  + 50a - 103, -29, 0, -37}}

\end{verbatim}}
    }}
\end{example}

\section{findANonZeroMinor, extendIdealByNonZeroMinor}

@@

\vspace{1em}
-- {\tt findANonZeroMinor(n, M, I)}

\vspace{1em}
\noindent where $I$ is an ideal
in a polynomial ring over {\tt QQ} or {\tt ZZ/p} for $p$ prime; $M$ is a matrix
over the polynomial ring and $n\in \mathbb{Z}$ denotes the size of the minors of interest.

@@

\item[\tt MinorPointAttempts => ZZ] 
  This controls how many points at which to check the rank of the matrix.
\end{description}

\vspace{1em}

\begin{example}~~

  ~~
  {{\small\color{blue}
\begin{verbatim}
		i3 : R = ZZ/5[x, y, z];

@@

		o6 : Sequence
\end{verbatim}}
    }}
\end{example}

@@

\vspace{1em}
-- {\tt extendIdealByNonZeroMinor(n, M, I)}

\vspace{1em}
\noindent where $n,M,I$ are same as before. This function finds a submatrix of size $n\times n$ using \\{\tt findANonZeroMinor};  
it extracts the last entry of the output, finds its determinant and
adds it to the ideal $I$, thus extending $I$.

% \vspace{1em}
As before, we can have the following additional options:
\begin{description}
\item[\tt Strategy => Symbol] This specifies which strategy to use when calling {\tt randomPoints}.
\item[\tt Homogeneous => Boolean] This is passed in calls to {\tt randomPoints}.
\item[\tt Verbose => Boolean] This turns on or off verbose output (see \Cref{findANonZeroMinor}).
\item[\tt MinorPointAttempts => ZZ] This controls how many points at which to check the rank of the matrix.
\end{description}

\vspace{1em}
\begin{example}~~

@@

o5 : Ideal of R
\end{verbatim}}	
    }}	
\end{example}

One use for this function can be in showing that certain rings are $(R_1)$ (regular in codimension $1$).
Consider the following example which is $(R_1)$ where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.  However, we can use this function to quickly find interesting minors.

\vspace{1em}

\begin{example}~~

  ~~
  {{\small\color{blue}
\begin{verbatim}
		i2 : T = ZZ/101[x1, x2, x3, x4, x5, x6, x7];

@@

\end{verbatim}}
    }}

\end{example}

In this particular example, there tend to be about $5$ associated primes when adding the first minor to $J$, and so one would expect about $5$ steps as each minor computed most likely will eliminate one of those primes.

>>>>

\documentclass[11pt]{amsart}
\usepackage{calc,amssymb,amsthm,amsmath,fullpage}
% \usepackage{mathtools}
\RequirePackage[dvipsnames,usenames, table]{xcolor}
\usepackage{hyperref}
\usepackage[capitalize]{cleveref}
\hypersetup{
  bookmarks,
  bookmarksdepth=3,
  bookmarksopen,
  bookmarksnumbered,
  pdfstartview=FitH,
  colorlinks,backref,hyperindex,
  linkcolor=Sepia,
  anchorcolor=BurntOrange,
  citecolor=MidnightBlue,
  citecolor=OliveGreen,
  filecolor=BlueViolet,
  menucolor=Yellow,
  urlcolor=OliveGreen
}
\usepackage{xypic}
\usepackage{alltt}
\usepackage{multicol}  
% \usepackage{etex}
\usepackage{xspace}
\usepackage{rotating}
\interfootnotelinepenalty=100000
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
%\usepackage{enumitem}
%\SetLabelAlign{parleft}{\parbox[t]\textwidth{#1\smallskip}}

% \usepackage{mabliautoref}
% \usepackage{colonequals}
% \frenchspacing
% \input{kmacros3.sty}
% \usepackage{stmaryrd}
% 
\usepackage{verbatim}
\usepackage{enumerate}

\begin{document}
\title{{RandomRationalPoints} package for \emph{Macaulay2}}
\author{Sankhaneel Bisui}
\email{sbisui@tulane.edu}
\address{Tulane University}
\author{Zhan Jiang}
\email{zoeng@umich.edu}
\address{University of Michigan}
\author{Sarasij Maitra}
\email{sm3vg@virginia.edu}
\address{University of Virginia}
\author{Thai Nguyen}
\email{tnguyen11@tulane.edu}
\address{Tulane University}
\author{Karl Schwede}
\email{schwede@math.utah.edu}
\address{Department of Mathematics, University of Utah, 155 S 1400 E Room 233, Salt Lake City, UT, 84112}
\date{\today}

\begin{abstract}
  In this article, we present {\tt RandomRationalPoints}, a package in \emph{Macaulay2} designed mainly to identify rational and geometric points in a variety over a finite field. We provide different strategies such as linear intersection and projection to obtain such points. We also present methods to obtain non-vanishing minors of some given size in a given matrix, by evaluating the matrix at a point.  %This, in conjunction with the {\tt FastLinAlg} package, should aid in determining properties of the singular locus.
\end{abstract}

@@

    There is an existing package called {\tt RandomPoints}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one point quickly, even if it is not rational.  
    We provide some functions that apply different strategies to generate random rational points for the given variety. We also provide functions that will expedite the process of determining properties of the singular locus of $X$.% when combined with the functions provided in the package {\tt FastLinAlg}.

    We provide the following core functions:
    \begin{itemize}    
    \item {\tt randomPoints}:  This tries to find a point in the vanishing set of an ideal. (\Cref{randomPoints})
    \item {\tt projectionToHypersurface} and {\tt genericProjection}: These functions provide customizable projection. (\Cref{projectionfunctions}) 	
    \item {\tt findANonZeroMinor} and {\tt extendIdealByNonZeroMinor}:  The first function finds a submatrix of a given matrix that is nonsingular at a point of a given ideal.  The second, adds said submatrix to an ideal, which is useful for computing partial Jacobian ideals.  (\Cref{findANonZeroMinor})          
    %\item :  This extends an ideal by a minor produced by \\{\tt findANonZeroMinor}. (\Cref{extendIdealByNonZeroMinor})	
    \end{itemize}

    All polynomial rings considered here will be over finite fields. In the subsequent sections, we explain briefly the core and some helper functions and describe the strategies that we have implemented in their execution.

@@

\section{Our primary pupose: {\tt randomPoints}}\label{randomPoints}
    We start with the core function in this package:  {\tt randomPoints} is a function to find rational or geometric points in a variety. The typical usages are as follows:

    \vspace{0.5em}
    -- {\tt randomPoints(I)},

@@

    \vspace{0.5em}
    \noindent where $n$ is a positive integer denoting the number of desired points, and 
    $I$ is an ideal inside a polynomial ring.  If {$ n$} is omitted, then it is assumed to be 1.

@@

    \begin{description}%[]
        \setlength{\itemsep}{5pt}
    \item[\tt Strategy => $\bullet$]   Here the $\bullet$ can be {\tt Default, BruteForce, LinearIntersection, \\ GenericProjection} or {\tt HybridProjectionIntersection}.

    \begin{itemize}
    \item {\tt Default} performs a sequence of the different strategies below, aimed at finding a point quickly.  It begins with less general projections and linear intersections, and gradually ramps up the randomness.  
    \item {\tt BruteForce} simply tries random points and sees if they are on the variety.

@@

    \begin{example}\label{BruteForce}
        Consider the following example.
        ~~        
        {\small\color{blue}
    \begin{quote}
\begin{verbatim}
i2 : R = ZZ/101[x, y, z];

@@

o6 : List
        \end{verbatim}
    \end{quote}
        }
    \end{example}%
    %
    \vspace{-1em}
    \item[\tt ProjectionAttempts => ZZ]

@@

    \item[\tt MaxCoordinatesToReplace => ZZ] This is used for fine tuning the {\tt LinearIntersection} or {\tt GenericProjection} strategies, see \Cref{helper}.
    \item[\tt Codimension => ZZ] If the codimension is already known, one can provide it so it is not recomputed.  This is used in the {\tt GenericProjection} and {\tt LinearIntersection} strategies, see \Cref{projectionfunctions}.

    \item[\tt ExtendField => Boolean]

@@

    \item[\tt IntersectionAttempts => ZZ]

@@

    \item[\tt PointCheckAttempts => ZZ]

    When calling {\tt randomPoints} with a {\tt BruteForce} strategy, this denotes the number of trials for brute force point checking.

    \begin{example}
        We re-compute \cref{BruteForce} this time specifying more attempts.
    {{\small\color{blue}
    ~~
    \begin{quote}
\begin{verbatim}    
i7 : randomPoints(J, Strategy => BruteForce, PointCheckAttempts => 20000)
    -- used 0.84679 seconds

o7 = {{-39, -43, 28}}
    \end{verbatim}%        
\end{quote}%
}}%
    \end{example}%
    \vspace{-1em}
%
    %\vspace{1em}
    \item[\tt MaxCoordinatesToTrivialize]

@@

%    \vspace{1em}
    \item[\tt NumThreadsToUse => ZZ]

    When calling {\tt randomPoints} with a {\tt BruteForce} strategy, this denotes the number of threads to use in brute force point checking.
    \end{description}
    % Outputs
    % :List
    % a list of points in the variety with possible repetitions.
    % Description
    % Text  
    % Gives at most $n$ many point in a variety $V(I)$. 
    % Example
    % R = ZZ/5[t_1..t_3];
    % I = ideal(t_1,t_2+t_3);
    % randomPoints(3, I)
    % randomPoints(4, I, Strategy => Default)
    % randomPoints(4, I, Strategy => GenericProjection)
    % randomPoints(4, I, Strategy => LinearIntersection)

@@

    When working over very small fields, frequently {\tt BruteForce} is most efficient.  This is not surprising as there may not be many points to check, and a surprisingly large percentage of points can be rational in hypersurfaces and complete intersections.  However, if the field size is larger, {\tt BruteForce} will perform poorly.  Other strategies work differently on different examples, and even the same strategy can sometimes work very quickly even if it typically works very slowly.  A quick comment, while in many examples generic projection works well, there are some where it has very slow performance.  Setting {\tt ProjectionAttempts=>0} in the {\tt Default} strategy is a way to avoid this situation.

    \begin{example}
        We begin with an example over a small field.
    {{\small\color{blue}
    ~~
    \begin{quote}
\begin{verbatim}
i2 : R = ZZ/7[x_1..x_10];

@@

o5 : List
    \end{verbatim}
\end{quote}\vspace{-1em}
    }}
    \end{example}

\begin{example}Now we work over a larger field.
        {{\small\color{blue}
        ~~
        \begin{quote}
\begin{verbatim}
i6 : S = ZZ/211[x_1..x_10];

@@

o11 = {{31, -15, 53, 41, 87, 9, -21, -43, -12, 85}}      
    \end{verbatim}
\end{quote}\vspace{-1em}
        }}
    \end{example}

    \begin{example}
        Finally, we allow our functions to extend our field.
        {{\small\color{blue}
        ~~
        \begin{quote}
        \begin{verbatim}
i12 : time randomPoints(J, Strategy => LinearIntersection, ExtendField => true)
-- used 0.0612306 seconds

            5      4      3      2                                      5  
o12 = {{54, 14a  - 48a  - 56a  + 56a  - 13a + 48, -101, 56, -42, -27, - 26a  -
    -------------------------------------------------------------------------
    4      3      2
    43a  - 70a  - 96a  + 50a - 103, -29, 0, -37}}
        \end{verbatim}
    \end{quote}
        }}
    \end{example}

\section{A helper function: {\tt randomCoordinateChange}}{\label{helper}}

% \item {\tt Codimension}: It takes input an integer. Various functions need to know the codimension/height of the scheme/ideal it is working with.  Setting this to be an integer will tell the function not to compute the codimension and to use this value instead.  The default value is {\tt null}, in which case the function will compute the codimension.  
%%%   
%%   Codimension is not a function, it is an option.
%%%   
\subsection*{\tt randomCoordinateChange:}

This function takes a polynomial ring as an input and outputs the coordinate change map.  
i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.

In some applications, a full change of coordinates is not desired, as it might cause code to run very slowly.  A binomial change of coordinates might be preferred, or we might only replace some monomials by other monomials.  
This is controlled with the following options.

\begin{itemize}
\item {\tt Replacement}: Setting {\tt Replacement => Full} will mean that coordinates are replaced by a general degree $1$ form.  If {\tt Replacement => Binomial}, the coordinates are only changed to binomials, which can be much faster for certain applications. If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
\item {\tt MaxCoordinatesToReplace}: The user can specify that only a specified number of coordinates should be non-monomial (assuming {\tt Homogeneous => true}).  
  This option is passed to {\tt randomCoordinateChange} by other functions that call it. 		
\item {\tt Homogeneous}:  Setting {\tt Homogeneous => false} will cause degree zero terms to be added to modified coordinates (including monomial coordinates).		
\end{itemize} 
%\par
%\vspace{-1em}    
\begin{example}  We demonstrate some of these options.
%
  {{\small\color{blue}
  \begin{quote}
\begin{verbatim}
	i3 : R = ZZ/11[x, y, z];

	o7 = | y x+4z z |
\end{verbatim}
\end{quote}
    }}

\end{example}
\section{Other helper functions: {\tt genericProjection}, {\tt projectionToHypersurface}}\label{projectionfunctions}

% \begin{enumerate}
\subsection{genericProjection} 
This function finds a random (somewhat, depending on options) generic projection of the ring or ideal.
The typical usages are as follows: 
\vspace{0.5em}

\vspace{0.5em}
\noindent where 
$I$ is an ideal 
in a polynomial ring, 
$R$ can denote a quotient of a polynomial ring and 
$n\in \mathbb{Z}$ is
an integer specifying how many dimensions to drop.  Note that this function makes no attempt to verify that the projection is actually generic with respect to the ideal.

\begin{example}	
    We project a curve in 4-space to 2-space.
    ~~
  {{\small\color{blue}
  \begin{quote}
\begin{verbatim}
i1 : R = ZZ/5[x, y, z, w];

o3 : Sequence
\end{verbatim}
\end{quote}
    }}
\end{example}

\begin{comment}~~

o3 : Sequence
\end{verbatim}}
    }}
\end{comment}

This method works by calling {\tt randomCoordinateChange} (see \Cref{helper}) before dropping some variables.  It passes the options {\tt Replacement}, {\tt MaxCoordinatesToReplace}, {\tt Homogeneous} to that function.
%The user can also specify the same options as before to be passed on to the {\tt randomCoordinateChange} function.

\begin{comment}~~

o4 : Sequence
\end{verbatim}}
    }}
\end{comment}

\vspace{0.5em}
-- {\tt projectionToHypersurface I},

-- {\tt projectionToHypersurface R} 
\vspace{0.5em}

\noindent where $I$ is an ideal in a polynomial ring, respectively, $R$ is a quotient of a polynomial ring. The output is
a list with two entries: the generic projection map and the ideal (respectively the ring).

%\vspace{1em}
\begin{comment}~~

	i3 : I = ideal(random(3, R) - 2, random(2, R));

	            ZZ                             4    3      2 2     4
	o4 = (map(R,--[y, z],{x + 2y, 2y}), ideal(y  + y z - 2y z  - 2z  - y - z))
	            5

	o5 : Sequence
	\end{verbatim}
    }}
\end{comment}

\section{An application:  {\tt findANonZeroMinor}, {\tt extendIdealByNonZeroMinor}}

@@

\vspace{0.5em}
-- {\tt findANonZeroMinor(n, M, I)} 
\vspace{0.5em}
\noindent where $I$ is an ideal
in a polynomial ring over {\tt QQ} or {\tt ZZ/p} for $p$ prime; $M$ is a matrix
over the polynomial ring and $n\in \mathbb{Z}$ denotes the size of the minors of interest.

@@

\item[\tt MinorPointAttempts => ZZ] 
  This controls how many points at which to check the rank of the matrix.
\end{description}
\vspace{0.25em}
\begin{example}
  We demonstrate this function.
  ~~
  {{\small\color{blue}
  \begin{quote}
\begin{verbatim}
		i3 : R = ZZ/5[x, y, z];

@@

		o6 : Sequence
\end{verbatim}
\end{quote}\vspace{-1em}    
    }}
\end{example}

@@

\vspace{0.5em}
-- {\tt extendIdealByNonZeroMinor(n, M, I)}

\vspace{0.5em}
\noindent where $n,M,I$ are same as before. This function finds a submatrix of size $n\times n$ using \\{\tt findANonZeroMinor};  
it extracts the last entry of the output, finds its determinant and
adds it to the ideal $I$, thus extending $I$.   It has the same options as {\tt findANonZeroMinor}.

% \vspace{1em}

%\vspace{1em}
\begin{comment}~~

@@

o5 : Ideal of R
\end{verbatim}}	
    }}	
\end{comment}

One can use this function to show that rings are $(R_1)$, that is, regular in codimension $1$.

\begin{example}
    Consider the following example (which is $(R_1)$) where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.  However, we can use this function to quickly find interesting minors.
  {{\small\color{blue}
  \begin{quote}
\begin{verbatim}
		i2 : T = ZZ/101[x1, x2, x3, x4, x5, x6, x7];

@@

\end{verbatim}
\end{quote}\vspace{-1em}    
    }}    
    \noindent
    In this particular example, there tend to be about $5$ associated primes when adding the first minor to $J$, and so one would expect about $5$ steps as each minor computed most likely will eliminate one of those primes.
\end{example}
